---
title: "Wk4 Processing Script"
author: "PCR"
date: "2023-07-19"
output: 
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(reshape2)
```
## Overview

This report presents preliminary data analysis for Week 3 Example Data. Specifically, this will report basic descriptiveness statisitcs as well as report tests of the following predictions: 

1. People will feel more guilt when:
  + 1.1 They caused the shock (vs. their partner)
  + 1.2 Their partner was morally good (vs. bad)
  + 1.3 The shock was stronger (high > medium > low)
  
2. Women will report more guilt than men
3. Agreeable pipel will report more guilt
4. Neurotic people will report more guilt

and

5. People willl be more generou swhen :
  + 5.1 They caused the shock (vs. their partner)
  + 5.2 Their partner was morally good (vs. bad)
  + 5.3 The shock was stronger (high > medium > low)
  
6. Women will be more generous guilt than men
7. Agreeable pipel will be more generous
8. Neurotic people will be more generous


```{r recode}
data <- read.csv("Week 3/Processed Data/Week 3 Data PROCESSED.csv")

##Recode variables:
#Lets recode Pin
data$PIN<-as.factor(data$PIN)

data$sex<-as.factor(data$sex)

data$shock <-as.factor(data$shock)

data$Who.did.it.<-as.factor(data$Who.did.it.)

data$pMoral<-as.factor(data$pMoral)
```

## Descriptives
Decrptive statistics for guild:
```{r descpGuilt}
c(summary(data$guilt),"SD"=sd(data$guilt,na.rm=T))
```

```{r descpGenerousity}
c(summary(data$preShare),"SD"=sd(data$preShare,na.rm=T))

c(summary(data$postShare),"SD"=sd(data$postShare,na.rm=T))
```

```{r Personality}
#Descriptives for agreeableness
c(summary(data$agree),"SD"=sd(data$agree,na.rm=T))

#Descriptives for neuroticsm
c(summary(data$neuro),"SD"=sd(data$neuro,na.rm=T))
```


## Guilt Analyses

### Prediction 1.1 
```{r p1.1}
#Run t-test for predictiion
p1.1 <- t.test(data$guilt~data$Who.did.it.)

#This means outcome ~ predictor
```
We found that 1.1 is supported. People who caused the shock felt more guilt than those who felt their partner caused the shock. 



### Prediction 1.2 
Prediction 1.2 People felt more guilt when their partner was morally good: 
```{r p1.2}
#We are doing a t-test. 

p1.2 <- t.test(data$guilt~data$pMoral)

#Output the results
p1.2


```


### Prediction 1.3
Prediction 1.3, people felt more guilt when their partner received a stronger shock. 

```{r}
p1.3 <- lm(guilt~shock, data=data)

#LM means linear model 
#Run anova where shock is outcome variable and guilt is predictor

#Output the results
anova(p1.3)

#Output the groups means
tapply(data$guilt,data$shock, function(x) mean(x,na.rm=T))
```


##Prediction 2

Prediction was not suppoprted. Women did not report more guilt than men:
```{r p2}
#Run the t-test for prediction 2
p2<- t.test(data$guilt~data$sex)

#Output the results
p2
```


##Prediction 3 - Agreeable people will report more guilt

Have to get this check, data is different
Prediction 3 was 

```{r p3}
#Runs the correlation test for predction 3
p3 <- cor.test(data$guilt,data$agree)

p3
```


##Prediction 4 - Neurotic people will report more guilt
Prediction was not supported. Neurotic were not more guilt.
```{r p4}
#Runs the correlation test for predction 3
p4 <- cor.test(data$guilt,data$neuro)

p4
```


##Generosity Analyses
```{r melt}
#Goin to melt it from shor to long

dataMelt<- melt(data,
                  measure.vars=c("preShare","postShare"),
                  variable.name=c("prePost"),
                  value.name=c("share")
                    )
                  
        

#This is going to create a new variable to keep track of what measure is which
```



##Prediction 5.1 - People willl be more generous if they caused the shock (vs. their partner)

Does seem that prePost, Shock and who caused the shock have a signficance. 
Preduction 5.1 
The data found that if the partner caused the shock they were less generous, while they were more generous if they themselves caused the shick. 
```{r p5.1}
#Run within subjects anova
p5.1 <- aov(share~prePost*Who.did.it.+Error(PIN),data = dataMelt)
#specificy an interaction is the use of *

summary(p5.1)

tapply(dataMelt$share,dataMelt[,c(24,21)], function(x) mean(x,na.rm=T))
```

